OptiFit: an improved method for fitting amplicon sequences to existing OTUs

### 2021-09-02 paper/paper.Rmd
Use 'improved' instead of 'robust' in some places

---
title: "OptiFit: a new method for fitting amplicon sequences to existing OTUs"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
output:
  pdf_document:
    includes:
      in_header: preamble.tex
      before_body: head.tex
    fig_caption: yes
csl: msystems.csl
fontsize: 12pt
geometry: margin=1.0in
repository_url: https://github.com/SchlossLab/OptiFitAnalysis
params:
    snakemake: "none"
---

@@

Assigning amplicon sequences to operational taxonomic units (OTUs) is often an
important step in characterizing the composition of microbial communities across
large datasets.
OptiClust, a _de novo_ OTU clustering method, has been
shown to produce higher quality OTU assignments than other methods and at
comparable or faster speeds.
A notable difference between _de novo_ clustering and database-dependent methods
is that OTU assignments from _de novo_ methods may change when new
sequences are added to a dataset.
However, in some cases one may wish to incorporate new samples into a previously
clustered dataset without performing clustering again on all sequences, such as
when comparing across datasets or deploying machine learning models where OTUs
are features.
**TODO**: sentence mentioning open/closed reference clustering and the inherent problems with the current methods.
To provide an efficient and robust method to fit amplicon sequence data to
existing OTUs, we developed the OptiFit algorithm.
We tested OptiFit using four microbiome datasets with two different strategies:
by fitting to an external reference database or by splitting the dataset into a 
reference and query set and fitting the query sequences to the reference set
after clustering it using OptiClust.
The result is a robust implementation of closed and open-reference clustering.
OptiFit produces OTUs of similar quality as OptiClust and at faster speeds when
using the split dataset strategy, although the OTU quality and processing speed
depends on the database chosen when using the external database strategy.
OptiFit provides a suitable option for users who require consistent OTU
assignments at the same quality afforded by _de novo_ clustering methods.

>>>>

---
title: "OptiFit: an improved method for fitting amplicon sequences to existing OTUs"
date: '`r Sys.Date()`'
bibliography: references.bib
link-citations: true
output:
  pdf_document:
    includes:
      in_header: preamble.tex
      before_body: head.tex
    fig_caption: yes
csl: msystems.csl
fontsize: 12pt
geometry: margin=1.0in
repository_url: https://github.com/SchlossLab/OptiFitAnalysis
params:
    snakemake: "none"
---

@@

Assigning amplicon sequences to operational taxonomic units (OTUs) is often an
important step in characterizing the composition of microbial communities across
large datasets.
OptiClust, a _de novo_ OTU clustering method, has been
shown to produce higher quality OTU assignments than other methods and at
comparable or faster speeds.
A notable difference between _de novo_ clustering and database-dependent methods
is that OTU assignments from _de novo_ methods may change when new
sequences are added to a dataset.
However, in some cases one may wish to incorporate new samples into a previously
clustered dataset without performing clustering again on all sequences, such as
when comparing across datasets or deploying machine learning models where OTUs
are features.
**TODO**: sentence mentioning open/closed reference clustering and the inherent problems with the current methods.
To provide an efficient and robust method to fit amplicon sequence data to
existing OTUs, we developed the OptiFit algorithm.
We tested OptiFit using four microbiome datasets with two different strategies:
by fitting to an external reference database or by splitting the dataset into a 
reference and query set and fitting the query sequences to the reference set
after clustering it using OptiClust.
The result is an improved implementation of closed and open-reference clustering.
OptiFit produces OTUs of similar quality as OptiClust and at faster speeds when
using the split dataset strategy, although the OTU quality and processing speed
depends on the database chosen when using the external database strategy.
OptiFit provides a suitable option for users who require consistent OTU
assignments at the same quality afforded by _de novo_ clustering methods.
